Enhanced proximity effect in zigzag-shaped Majorana Josephson junctions.

### 2019-02-20 paper/snakemajoranas.tex
small review with Anton

\comment{We consider a two-dimensional system with zigzag and BdG Hamiltonian.}
We consider a Josephson junction (Fig.~\ref{fig:setup}) consisting of a 2D strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either saw-toothed zigzag as depicted [Fig.~\ref{fig:setup}(b)], or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{equation}
\begin{split}
    H_\textrm{N} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z
        + E_\text{z} \sigma_x
        + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \tau_z \\
    \\
    H_\textrm{SC} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z
        + \Delta \cos{\frac{\phi}{2}} \tau_x + \Delta \sin{\frac{\phi}{2}} \tau_y.
\end{split}
\label{eq:hamiltonian}
\end{equation}
Where $H_\textrm{N}$ and $H_\textrm{SC}$ are the Hamiltonians of the semiconductor and superconductors, respectively.
They act on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction [see Fig.~\ref{fig:setup}(b)].
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.

@@

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
The resulting Hamiltonians are large because instead of having a single slice of sites that is translationally invariant in the $x$-direction, we create a long 2D supercell to make the structure periodic.
To speed up these expensive simulations we employ an adaptive sampling strategy, using the Python package Adaptive.~\cite{adaptive}
The entire source code and the resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/bandstructures}
\caption{Figure of the band stuctures corresponding to the system in Fig.~\ref{fig:setup} with different zigzag amplitudes.
The blue lines correspond to $\phi=0$, $B=0$ (trivial phase) and the orange lines to $\phi=\pi$, $B = \SI{1}{T}$ (topological phase).
The three subplots are for different amplitudes of the zigzag, with (a) a straight system $z_y=0$, (b) $z_y=\frac{W}{4}$, and (c) $z_y=\frac{W}{2}$, where $W=\SI{200}{\nm}$ is the junction width.
Subplot (a) has a different scale on the $x$-axis for $k_x < 0$ than the other subplots and displays the unfolded band structure.
For the right-hand side of (a) ($k_x > 0$), (b), and (c), the folding is the same, such that the velocity $v=\frac{dE}{dk}$ can be compared visually.
We observe that once there are no more straight trajectories inside the junction (when $z_y=\frac{W}{2}$) the spectrum becomes most insensitive to the momentum $k_x$, and thus the $v_\textrm{F}$ decreases.
As the zigzag amplitude increases the band gap $E_\textrm{gap}$ increases by an order of magnitude.
The combination of these ensures a significant decrease of the Majorana decay length because $\xi_M \propto \frac{v_\textrm{F}}{E_\textrm{gap}}$
The parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:band_stuctures}}
% TODO: mention z_x!
\end{figure}

\comment{We calculate the band structure for varying amount of zigzag.}
In figure~\ref{fig:band_stuctures}, the bandstructure for zigzag systems with varying amplitude is displayed.
Because the discretized zigzag system is modeled using a supercell, we can only plot a folded bandstructure.
For comparison's sake, we take the same cell size for the ribbon structure such that the bandstructure folds in the same way.

\comment{The band structures show that modulation of the geometry increases the gap by an order of magnitude, as well as reduce the group velocity.}
The introduction of a zigzag has a striking effect: the bands flatten out because of a reduction in velocity and the topological gap increases by more than an order of magnitude.
On an intuitive level we can understand this because in a straight system long trajectories are possible, these have a momentum that is directed along the strip and consequently, $k_x \approx k_F$.
Further, the smallest gap in a straight system occurs around $k_F$, and because the zigzag cuts off long trajectories, the gap size increases.
Besides this, the velocity of the modes greatly reduces, which has a positive effect on the Majorana coherence length, as described later in section \ref{sec:shape_effects}.

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3L_x=\SI{3900}{\nm}$, the remaining parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:wavefunctions}}
\end{figure}

\comment{We calculate the wave functions and find the Majorana lengths by fitting an exponential.}
We model a finite system to compute the Majorana wave function density for different geometries; ribbon, saw-toothed zigzag, parallel curves, and variants of the latter with disordered edges.
By diagonalizing the Hamiltonian, we find the Majorana energy $E_\textrm{M}$ and the energy of the first excited state $E_1$, the lowest and second lowest energies, respectively.
Using the eigenstates, we calculate the Majorana decay length $\xi_\textrm{M}$ by fitting an exponential to the density of a single Majorana wavefunction projected on the $x$-axis.
To determine $\xi_\textrm{M}$, we artificially break the particle-hole symmetry on one edge by adding a potential $V \sigma_0 \tau_0$ such that one of the Majorana states is pushed into the continuum and the fit to $\exp{\frac{x}{\xi_\textrm{M}}}$ becomes easier.
In figure~\ref{fig:wavefunctions} the wavefunctions are displayed, superimposed upon their respective geometry.
All systems have identical Hamiltonian parameters in order to compare the effects of the changing geometry meaningfully.

>>>>

\comment{We consider a two-dimensional system with zigzag and BdG Hamiltonian.}
We consider a Josephson junction (Fig.~\ref{fig:setup}) consisting of a 2D strip of semiconductor, with superconductors on both sides.
We modulate the shape of the normal region, which can be either saw-toothed zigzag as depicted [Fig.~\ref{fig:setup}(b)], or a more smooth sinusoidal-like shape.
Similar to the conventional straight system \cite{pientka2017topological}, a magnetic field $B_x$ perpendicular to the junction is applied.
We model the system with a Bogoliubov-de Gennes Hamiltonian (BdG) Eq.~\eqref{eq:hamiltonian}; the normal part has a linear Rashba spin-orbit coupling term characterized by $\alpha$ and a Zeeman field with $E_z=\frac{1}{2} \mu_B g B_x$.
The superconductor has a superconducting coupling term $\Delta$ with the pair of superconductors at an opposite superconducting phase $\exp{\pm i \phi/2}$.
Both the normal part and the superconducters have a kinetic term and chemical potential $\mu$.
The complete Hamiltonian has the form
\begin{subequations}
\begin{align}
    H_\textrm{N} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu + \alpha \left( \ky \sigma_x - \kx \sigma_y \right) \right] \tau_z
        + E_\text{z} \sigma_x \\
    H_\textrm{SC} = & \left[\frac{\hbar^2\left(\kx^2 + \ky^2\right)}{2\meff} - \mu\right] \tau_z
        + \Delta \cos{\frac{\phi}{2}} \tau_x + \Delta \sin{\frac{\phi}{2}} \tau_y.
\end{align}
\label{eq:hamiltonian}
\end{subequations}
Where $H_\textrm{N}$ and $H_\textrm{SC}$ are the Hamiltonians of the semiconductor and superconductors, respectively.
They act on the spinor wave function $\Psi={\left(\psi_{e\uparrow},\psi_{e\downarrow},\psi_{\textrm{h}\downarrow},-\psi_{\textrm{h}\uparrow}\right)}^{T}$, where $\psi_e$, $\psi_\textrm{h}$ are its electron and hole components, and $\psi_\uparrow$, $\psi_\downarrow$ are the spin-up and spin-down components.
The Pauli matrices $\sigma_{i}$ act on the spin degree of freedom and $\tau_{i}$ act on the electron-hole degree of freedom.
For the zigzag geometry, we first consider a sawtooth-like pattern where $z_x$ is periodicity, $z_y$ the amplitude and $W$ the width of the junction [see Fig.~\ref{fig:setup}(b)].
Later in this manuscript, we show that the exact details of the shape are unimportant and consider different zigzag geometries also parametrized by $z_x, \; z_y, \; W$.

\comment{We discretize the Hamiltonian and simulate it with Kwant.}
We discretize our continuum Hamiltonian [Eq.~\eqref{eq:hamiltonian}] on a square grid and implement a tight-binding model using Kwant~\cite{groth_kwant:_2014}.
To optimize the sampling in important regions of parameter space, we use the Adaptive package.~\cite{adaptive}
The entire source code and the resulting raw data are available in Ref.~\onlinecite{data}.  % TODO: update the bib when the data is uploaded

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/bandstructures}
\caption{Figure of the band stuctures corresponding to the system in Fig.~\ref{fig:setup} with different zigzag amplitudes.
The blue lines correspond to $\phi=0$, $B=0$ (trivial phase) and the orange lines to $\phi=\pi$, $B = \SI{1}{T}$ (topological phase).
The three subplots are for different amplitudes of the zigzag, with (a) a straight system $z_y=0$, (b) $z_y=\frac{W}{2}$, and (c) $z_y=W$, where $W=\SI{200}{\nm}$ is the junction width.
On the $x$-axis is the folded lattice momentum where $z_x=\SI{1300}{\nm}$.
Subplot (a) has a different scale on the $x$-axis for $k_x < 0$ than the other subplots and displays the unfolded band structure.
For the right-hand side of (a) ($k_x > 0$), (b), and (c), the folding is the same, such that the velocity $v=\frac{dE}{dk}$ can be compared visually.
We observe that once there are no more straight trajectories inside the junction (when $z_y=W$) the spectrum becomes most insensitive to the momentum $k_x$, and thus the $v_\textrm{F}$ decreases.
As the zigzag amplitude increases the band gap $E_\textrm{gap}$ increases by an order of magnitude.
The combination of these ensures a significant decrease of the Majorana decay length because $\xi_M \propto \frac{v_\textrm{F}}{E_\textrm{gap}}$
The parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:band_stuctures}}
% TODO: mention z_x!
\end{figure}

\comment{We calculate the band structure for varying amount of zigzag.}
We diagonalize the supercell Hamiltonian to compute the band structure.
Because of the large periodicity of the zigzag and the resulting large supercell, the band structure is heavily folded.
In figure~\ref{fig:band_stuctures} we show the resulting band structures for zigzag systems with a varying amplitude.

\comment{The band structures show that modulation of the geometry increases the gap by an order of magnitude, as well as reduce the group velocity.}
The introduction of a zigzag has a striking effect: the bands flatten out and the topological gap increases by more than an order of magnitude.
On an intuitive level we can understand this because in a straight system long trajectories are possible, these have a momentum that is directed along the strip and consequently, $k_x \approx k_F$.
Further, the smallest gap in a straight system occurs around $k_F$, and because the zigzag cuts off long trajectories, the gap size increases.
Besides this, the states that exist in different segments have very little overlap and therefore have a very small velocity.
This reduction in velocity has a positive effect on the Majorana coherence length, as described later in section \ref{sec:shape_effects}.

@@

\begin{figure}[!htb]
\includegraphics[width=0.95\columnwidth]{figures/wavefunctions}
\caption{Density of wavefunctions $\left|\psi\right|^2$ for sizes and geometries.
With (a) a straight system, (b) a saw-toothed zigzag system, (c) a system where the normal region is defined by lines parallel to a sinusoid, and (d) similar to (c) but with disordered edges.
Inside the figure, we indicate the Majorana length (or coherence length) $\xi_\textrm{M}$, the Majorana energy $E_\textrm{M}$ (the energy of the first excited state), and the topological energy gap $E_\textrm{gap}$.
We observe that $\xi_\textrm{M}$ for (a) is almost two orders of magnitude longer and $E_\textrm{gap}$ more than an order of magnitude smaller than for (b), (c), and (d).
The robustness of $E_\textrm{gap}$ and $\xi_\textrm{M}$ across (b), (c), and (d) means that the details of the geometry do not matter for the improvements to occur.
The length of the system is $3.5 L_x=\SI{4550}{\nm}$, the remaining parameters values are written at the bottom of Sec.~\ref{sec:setup}.\label{fig:wavefunctions}}
\end{figure}

\comment{We calculate the wave functions and find the Majorana lengths by fitting an exponential.}
We model a finite system to compute the Majorana wave function density for different geometries; ribbon, saw-toothed zigzag, parallel curves, and variants of the latter with disordered edges.
By diagonalizing the Hamiltonian, we find the Majorana energy $E_\textrm{M}$ and the energy of the first excited state $E_1$, the lowest and second lowest energies, respectively.
Using the eigenstates, we calculate the Majorana decay length $\xi_\textrm{M}$ by fitting an exponential to the density of a single Majorana wavefunction projected on the $x$-axis.
To determine $\xi_\textrm{M}$, we use a trick and artificially break the particle-hole symmetry on one edge by adding a potential $V \sigma_0 \tau_0$ such that one of the Majorana states is pushed into the continuum and the fit to $\exp{\frac{x}{\xi_\textrm{M}}}$ becomes easier.
In figure~\ref{fig:wavefunctions} the wavefunctions are displayed, superimposed upon their respective geometry.
All systems have identical Hamiltonian parameters in order to compare the effects of the changing geometry meaningfully.
